---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gitlabr

<!-- badges: start -->
[![CRAN\_Status\_Badge](https://www.r-pkg.org/badges/version/gitlabr)](https://cran.r-project.org/package=gitlabr)
![CRAN\ Downloads\ Badge](https://cranlogs.r-pkg.org/badges/gitlabr)
[![R build status](https://github.com/statnmap/gitlabr/workflows/R-CMD-check/badge.svg)](https://github.com/statnmap/gitlabr/actions)
<!-- badges: end -->


## Installation

You can install the most recent stable version from CRAN using:

```{r, eval=FALSE}
install.packages("gitlabr")
```

To install the development version using [devtools](https://cran.r-project.org/package=devtools)), type:
```{r, eval=FALSE}
devtools::install_github("statnmap/gitlabr")
```

See the [CONTRIBUTING.md](CONTRIBUTING.md) for instructions on how to run tests locally and contributor information.

## Recommended Gitlab versions

Gitlab 11.6 or higher is generally recommended when using gitlabr version 1.1.6 or higher. This {gitlabr} version uses the gitlab API v4.

## Quick Start Example

R code using {gitlabr} to perform some common gitlab actions can look like this

- Store your token in .Renviron with `usethis::edit_r_environ()` and restart your session 

- Set a connection to Gitlab instance
```{r}
library(gitlabr)

# Add: GITLAB_COM_TOKEN=YourTokenHere
# You can verify it worked
Sys.getenv("GITLAB_COM_TOKEN")

# connect as a fixed user to a gitlab instance
my_gitlab <- gl_connection(
  gitlab_url = "https://gitlab.com",
  private_token = Sys.getenv("GITLAB_COM_TOKEN"))
# a function is returned
# its first argument is the request (name or function), optionally followed by parameters
```

- Find the list of projects available to you
  + Define a limit of pages of projects to search in with `max_page`, otherwise the entire Gitlab.com will be downloaded here...
```{r}
# a data_frame is returned, as is always by gitlabr functions
my_gitlab(gl_list_projects, max_page = 2) 
```

- Explore one of your projects. You can set the name of the project or its ID. The ID is highly recommended, in particular if your project does not appear in the first pages of projects above. 
  + Let's explore [project "repo.rtask"](https://gitlab.com/statnmap/repo.rtask), with `ID = 20384533` on Gitlab.com
  
```{r}
my_project <- 20384533 #repo.rtask",
```

- List files of the project using `gl_list_files`
```{r}
my_gitlab(gl_list_files, project = my_project)
```

- List issues with `gl_list_issues`

```{r}
my_gitlab(gl_list_issues, project = my_project)
```

- Create an issue

```{r}
# create a new issue
new_feature_issue <- my_gitlab(gl_new_issue, project = my_project, "Implement new feature")

# statnmap user ID
my_id <- 4809823

# assign issue to me
my_gitlab(gl_assign_issue, 
          project = my_project,
          new_feature_issue$iid,
          assignee_id = my_id)

# Verify new issue is here
my_gitlab(gl_list_issues, my_project, state = "opened")

# close issue
my_gitlab(gl_close_issue, project = my_project, new_feature_issue$iid)$state
```

## Further information

- For a comprehensive overview & introduction see the `vignette("quick-start-gitlabr")`
- When writing custom extensions ("convenience functions") for gitlabr or when you experience any trouble, the very extensive [Gitlab API documentation](http://doc.gitlab.com/ce/api/) can be helpful.

_Note that the {gitlabr} package was originally created by [Jirka Lewandowski](https://github.com/jirkalewandowski/gitlabr). The present repository is a fork to be able to continue development of this package._  


## To Do - If you want to contribute

*See the [CONTRIBUTING.md](CONTRIBUTING.md) for instructions on how to run tests locally and contributor information.*  
**Please help me pass unit tests !**

From [Jenny Brian review](https://github.com/jennybc/gitlabr/blob/jenny-review/jenny-review.md)

### Example with gitlab.com

I have not been able to run basic code against gitlab.com. I *can* run essentially all code I see in tests, vignette, etc. against gitlab.points-of-interest.cc. It's not clear if this is a real problem or a documentation problem, leading to user error. Either way, it needs to be addressed.
  + It would probably be best if examples, README, etc. featured gitlab.com, as most users would probably be doing that. It might also reveal now or in the future if there is some underlying problem that I am experiencing.
  + For example, here's code adapted from the examples in `gl_connection()`:
```r
my_gitlab <- gl_connection(
  "https://gitlab.com",
  private_token = Sys.getenv("GITLAB_PVT_TOKEN")
)
my_gitlab("projects")
```

  + First, the `gl_connection()` call needs to specify the name of the second argument or else it's interpreted as the `login`. Second, the `my_gitlab()` call hangs indefinitely for me. 
  
### Automated tests

*See the [CONTRIBUTING.md](CONTRIBUTING.md) for instructions on how to run tests locally and contributor information.*

I was able to run tests for this package without too much pain, which is pretty impressive, so big kudos for that! I have yet to write tests for an API-wrapping package that achieves this worthy goal. Some thoughts and suggestions:

  * Why test against your gitlab server vs gitlab.com? It's not completely hard-wired, given the URL in `environment.yml` but I suspect there would be some undocumented setup required if I actually changed the target gitlab server.
  * The instructions don't tell me how to populate `environment.yml`. Where do I get this info on gitlab? Is the token supposed to be the private token displayed in the account tab of user settings or an access token generated from the access tokens tab?
  * Once I populate `environment.yml` with real info, I should `.gitignore` it, right? That would be good to clarify. If testing on CI, then I assume there's an encryption strategy for this file. Or maybe secure env vars are used?
  * Each test file should use `context("blah blah")` at the top for nicer reporting. I've done this is my review branch.
  * I had to `skip()` the test about issue editing, because I don't have necessary permission. Would be good to implement this such that tester is editing an issue they have permission on. Again, pretty impressive that it's the *only* test I had to skip!  I've done this is my review branch.
  * [These lines re: API version in `testthat.R`](https://github.com/jirkalewandowski/gitlabr/blob/3525479b1316abd2f2122f7778be2c3c4f8b3d06/tests/testthat.R#L4-L14) are pretty unusual. I can't infer the exact goal, but wonder if there's a better / more conventional way to achieve it. Obviously, I understand it's got something to do with testing against different API versions.
  * I've moved common env var setup into `helper.R` and out of individual test files. Looking at the top of these files, I suspect even more setup could probably be similarly centralized? `helper.R` is always sourced by `devtools::load_all()` so this makes for a nice workflow even when developing / running tests interactively, i.e. the info doesn't *need* to live in each file.

# Contributing to {gitlabr}

You're welcome to contribute to {gitlabr} by editing the source code, adding more convenience functions, filing issues, etc. [CONTRIBUTING.md](CONTRIBUTING.md) compiles some information helpful in that process.

Please also note the [Code of Conduct](CONDUCT.md).
