---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = Sys.getenv("GITLAB_COM_TOKEN", unset = FALSE) # do not eval if token is not available
)
```

# {gitlabr}

<!-- badges: start -->
[![CRAN\_Status\_Badge](https://www.r-pkg.org/badges/version/gitlabr)](https://cran.r-project.org/package=gitlabr)
![CRAN\ Downloads\ Badge](https://cranlogs.r-pkg.org/badges/gitlabr)
[![R build status](https://github.com/statnmap/gitlabr/workflows/R-CMD-check/badge.svg)](https://github.com/statnmap/gitlabr/actions)
[![R-CMD-check](https://github.com/statnmap/gitlabr/workflows/R-CMD-check/badge.svg)](https://github.com/statnmap/gitlabr/actions)
<!-- badges: end -->

**There are multiple breaking changes in {gitlabr} v2, please refer to the corresponding vignette: https://statnmap.github.io/gitlabr/articles/gitlabr-v2.html**

_Note that the {gitlabr} package was originally created by [Jirka Lewandowski](https://github.com/jirkalewandowski/gitlabr). The present repository is a fork to be able to continue development of this package._  

## Installation

You can install the most recent stable version from CRAN using:

```{r, eval=FALSE}
install.packages("gitlabr")
```

To install the development version using [devtools](https://cran.r-project.org/package=devtools), type:
```{r, eval=FALSE}
devtools::install_github("statnmap/gitlabr")
```

See the [CONTRIBUTING.md](CONTRIBUTING.md) for instructions on how to run tests locally and contributor information.

## Recommended GitLab versions

GitLab 11.6 or higher is generally recommended when using {gitlabr} version 1.1.6 or higher. This {gitlabr} version uses the GitLab API v4.

## Quick Start Example

R code using {gitlabr} to perform some common GitLab actions can look like this

- Create a TOKEN on your Gitlab instance with scopes: `api`

- Store your token in .Renviron as `GITLAB_COM_TOKEN` with `usethis::edit_r_environ()` and restart your session 

- Set a connection to GitLab instance
```{r}
library(gitlabr)

# You can verify your token works
# Sys.getenv("GITLAB_COM_TOKEN")

# connect as a fixed user to a GitLab instance for the session
set_gitlab_connection(
  gitlab_url = "https://gitlab.com",
  private_token = Sys.getenv("GITLAB_COM_TOKEN"))
```

- Find the list of projects available to you
  + Define a limit of pages of projects to search in with `max_page`, otherwise the entire GitLab.com will be downloaded here...
```{r}
# a tibble is returned, as is always by {gitlabr} functions
gl_list_projects(max_page = 2) 
```

### Work with a specific project

- Explore one of your projects. You can set the name of the project or its ID. The ID is highly recommended, in particular if your project does not appear in the first pages of projects above. 
  + Let's explore [project "repo.rtask"](https://gitlab.com/statnmap/repo.rtask), with `ID = 20384533` on GitLab.com
  
```{r}
my_project <- 20384533 #repo.rtask",
```

- List files of the project using `gl_list_files`
```{r}
gl_list_files(project = my_project)
```

- List issues with `gl_list_issues`

```{r}
gl_list_issues(project = my_project)
```

- Create an issue

```{r}
# create a new issue
new_feature_issue <- gl_new_issue(project = my_project, title = "Implement new feature")

# statnmap user ID
my_id <- 4809823

# assign issue to me
gl_assign_issue(project = my_project,
                issue_id = new_feature_issue$iid,
                assignee_id = my_id)

# Verify new issue is here
gl_list_issues(project = my_project, state = "opened")

# close issue
gl_close_issue(project = my_project, issue_id = new_feature_issue$iid)$state
```

### Use additionnal requests

If an API request is not already available in {gitlabr}, function `gitlab()` allows to use any request of the GitLab API [https://docs.gitlab.com/ce/api/].

For instance, the API documentation shows how to create a new project in [https://docs.gitlab.com/ee/api/projects.html#create-project]:

- The verb is `POST`
- The request is `projects`
- Required attributes are `name` or `path` (if `name` not set)
- `default_branch` is an attribute that can be set if wanted, but not required

The corresponding use of `gitlab()` is:

```{r, eval=FALSE}
gitlab(
  req = "projects",
  verb = httr::POST,
  name = "toto",
  default_branch = "main"
)
```

Implement whatever suits your needs !

### Unset connection

```{r}
unset_gitlab_connection()
```

## Further information

- For a comprehensive overview & introduction see the `vignette("quick-start-guide-to-gitlabr")`
- When writing custom extensions ("convenience functions") for {gitlabr} or when you experience any trouble, the very extensive [GitLab API documentation](http://doc.gitlab.com/ce/api/) can be helpful.

# Contributing to {gitlabr}

You're welcome to contribute to {gitlabr} by editing the source code, adding more convenience functions, filing issues, etc. [CONTRIBUTING.md](CONTRIBUTING.md) compiles some information helpful in that process.

Please also note the [Code of Conduct](CONDUCT.md).
