---
title: "Quick Start Guide to gitlabr"
author: "Jirka Lewandowski <jirka.lewandowski@wzb.eu>"
date: "September 4, 2015"
output: pdf_document
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Quick Start Guide to gitlabr}
-->


# Quick Start Example

Here is example code for some very common tasks and to give an impression of how
to use `gitlabr`:

```{r eval=FALSE}
library(gitlabr)

# connect as a fixed user to a gitlab instance
my_gitlab <- gitlab_connection("http://gitlab.points-of-interest.cc",
                               login = "my_username",
                               password = readLines("secrets/gitlab_password.txt"))
# a function is returned
# its first argument is the request followed by parameters
my_gitlab(list_projects)
my_gitlab(list_files, "test-project", path = "subdir")

new_feature_issue <- my_gitlab(new_issue, "test-project", "Implement new feature")

# requests via gitlabr always return data.frames  ## TODO implement this

example_user <-
  my_gitlab("users") %>%
    filter(username == "testibaer")

my_gitlab(assign_issue, "test-project",
          new_feature_issue$iid,
          assignee_id = example_user$id)
```

- Common queries are each wrapped in special convenience functions (find a list in the section ["Function list"](#function-list)).
- `gitlabr` provides access the complete gitlab API - learn how to do this in the section ["Principles"](#principles). 
- Using `project_connection` instead of `gitlab_connection` also allows you to direct your queries to a specific project and get rid of the redundant repetition of the project name as argument - learn how to do this in the section ["Different ways to do it"](#different-ways-to-do-it).
- For a brief tour of the design principles of `gitlabr` refer to the section ["Principles"](#principles).
- Learn how to write your own convenience wrappers on top of the `gitlabr` logic in the section ["Writing custom gitlab request functions"](#Writing-custom-gitlab-request-functions).

# Philosophy

## Principles

### `gitlabr` provides a high and a low level interface at the same time.

While the low level interface allows to issue all possible queries to a gitlab API,
many common queries are simplified and wrapped in high level functions.

The core function of the **low level** interface is `gitlab`. It takes the request location as a character vector, API endpoint URL, HTTP verb as required arguments and passes additional arguments as query parameters (keeping their names) on to the API request.

```{r eval = FALSE}
gitlab(c("project", 12, "issues"), 
       api_root = "http://gitlab.points-of-interest.cc/api/v3",
       private_token = "XXX", # authentication for API
       verb = httr::GET,  # defaults to GET, but POST, PUT, DELETE can be used likewise
       state = "active")
```

translates to

```
GET http://gitlab.points-of-interest.cc/api/v3/project/12/issues?state=active&private_token=XXX
```

This way, any request documented in the [Gitlab API documentation](doc.gitlab.com/ce/api) can be issued from `gitlabr`.

The **high level** interface is consists of a number of functions that each have additional
arguments from which the request location is constructed, while all other arguments are simply passed on to `gitlab`. For example:

```{r eval = FALSE}
edit_issue(project = "test-project", 12, description = "Really cool new feature") ## TODO further arguments
```

does nothing but

```{r eval = FALSE}
gitlab(...) ## TODO
```

and hence translates to

```
GET TODO
```

Find a list of all implemented high level functions in the section ["Function list"](#function-list)) or learn how to write your own in the section ["Writing custom gitlab request functions"](#Writing-custom-gitlab-request-functions).

### `gitlabr` is implemented in the functional paradigm

passing a function as a request to `gitlab`

creating connections -> link to below

### Every output of a `gitlabr` call is a data.frame

TODO

### Pagination is hidden

TODO


## Different ways to do it

- two equivalent idioms: `gitlab(edit_issue, ...)` vs. `edit_issue(...)`
- creating connections

# Function list

Table: gitlabr name, API location, parameter translation (in parenthesis those documented by gitlabr but passed on without change)

Note: There are more locations and actions that can be accessed through the gitlab API. See the documentation of the Gitlab API for this.

# Writing custom gitlab request functions

- Link back to philosophy section
- remind that contribution is possible


