---
title: "Breaking changes in {gitlabr} V2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gitlabr-v2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gitlabr)
```

* Default branch is named `main` whenever required.
  + This can be changed with `gitlabr_options_set("gitlabr.main", "master")`
* All project-specific functions get `project` as first parameter. Changes apply to:
  + `gl_get_comments()`,`gl_to_issue_id()`, `gl_get_issue_comments()`, `gl_get_commit_comments()`,
  `gl_edit_comment()`, `gl_edit_issue_comment()`, `gl_edit_commit_comment()`
  + `gl_repository()`
  + `gl_get_issue()`, `gl_to_issue_id()`, `gl_new_issue()`, `gl_create_issue()`, `gl_edit_issue()`, `gl_close_issue()`,
  `gl_reopen_issue()`, `gl_assign_issue()`, `gl_unassign_issue()`
* Changed use of `api_version = "v4"` by `api_version = 4`
* Changed use of `force_api_v3 = TRUE` by `api_version = 4` for deprecation by default
